Efficiently Handling Guards in the Low Level Design of RPython's tracing JIT

### 2012-08-01 talk/vmil2012/paper.tex
extend contribution and structure paragraphs of the introduction

\begin{figure*}
    \include{figures/ops_count_table}
    \caption{Relative numbers of operations in the traces generated for
    different benchmarks}
    \label{fig:ops_count}
\end{figure*}
Although there are several publications about tracing just-in-time compilers, to
our knowledge, there are none that describe the use and implementation of
guards in this context. With the following contributions we aim to shed some
light (to much?) on this topic.
The contributions of this paper are:
\todo{more motivation}
Based on the informal observation that guards are among the most common
operations in the traces produced by PyPy's tracing JIT and that guards are
operations that are associated with an overhead to maintain information about
state to be able to rebuild it, our goal is to present concrete numbers for the
frequency and the overhead produced by guards, explain how they are implemented
in the different levels of PyPy's tracing JIT and explain the rationale behind
the design decisions based on the numbers.
As can be seen on Figure~\ref{fig:ops_count} guards account for 14.42\% to
22.32\% of the operations before and for 15.2\% to 20.12\% of after the
optimization pass over the traced and compiled paths of the benchmarks.
Figure~\ref{fig:benchmarks} shows the absolute number of operations for each
benchmark, for every guard that stays alive after optimization there are
several kinds of metadata created and stored at different levels of the JIT to
be able to rebuild the interpreter or tracer state from a guard failure making
the optimization \bivab{some good word} of guards an important aspect of the
low-level design of a tracing just-in-time compiler.
\todo{extend}
\todo{contributions, description of PyPy's guard architecture, analysis on benchmarks}
\begin{itemize}
 \item
\end{itemize}

The paper is structured as follows:

@@

\section{Resume Data}
\label{sec:Resume Data}

@@

\section{Related Work}

@@

% subsection Deoptimization in Method-Based JITs (end)

\section{Conclusion}

>>>>

\begin{figure*}
    \include{figures/ops_count_table}
    \caption{Relative numbers of operations in the traces generated for
    different benchmarks}
    \label{fig:ops_count}
\end{figure*}
Although there are several publications about tracing just-in-time compilers, to
our knowledge, there are none that describe the use and implementation of
guards in this context. With the following contributions we aim to shed some
light (to much?) on this topic.
The contributions of this paper are:
\todo{more motivation}
Based on the informal observation that guards are among the most common
operations in the traces produced by PyPy's tracing JIT and that guards are
operations that are associated with an overhead to maintain information about
state to be able to rebuild it, our goal is to present concrete numbers for the
frequency and the overhead produced by guards, explain how they are implemented
in the different levels of PyPy's tracing JIT and explain the rationale behind
the design decisions based on the numbers.
As can be seen on Figure~\ref{fig:ops_count} guards account for 14.42\% to
22.32\% of the operations before and for 15.2\% to 20.12\% of after the
optimization pass over the traced and compiled paths of the benchmarks.
Figure~\ref{fig:benchmarks} shows the absolute number of operations for each
benchmark, for every guard that stays alive after optimization there are
several kinds of metadata created and stored at different levels of the JIT to
be able to rebuild the interpreter or tracer state from a guard failure making
the optimization \bivab{some good word} of guards an important aspect of the
low-level design of a tracing just-in-time compiler.
The contributions of this paper are
In this paper we want to substantiate the aforementioned and describe based on
them the reasoning behind and the implementation of guards in PyPy's tracing
just-in-time compiler.
\begin{itemize}
  \item An analysis of guards in the context of PyPy's tracing JIT to
  substantiate the aforementioned observation, based on a set of benchmarks.
  \item We provide a detailed measurements about the frequency and the
  overhead associated with guards.
  \item We provide a description about how guards are implemented in the high\-
  and low-level parts of the JIT and describe the rationale behind the design.
\end{itemize}

The set of central concepts upon which this work is based is described in
Section~\ref{sec:Background}, such as the PyPy project, the RPython language
and its meta-tracing JIT. Based on these concepts in Section~\ref{sec:Resume
Data} we proceed to describe for PyPy's tracing JIT the details of guards in
the frontend\bivab{better term for this?} related to recording and storing the
information required to restore the interpreter state in case of a guard
failure, once the frontend has traced and optimized a loop it invokes the
backend to compile the operations to machine code, Section \ref{sec:Guards in
the Backend} describes the low-level aspects of how guards are implemented in
the JIT-backend. The frequency of guards and the overhead associated with the
implementation described in this paper is discussed in
Section~\ref{sec:evaluation}. Section~\ref{sec:Related Work} presents an
overview about how guards are treated in the context of other just-in-time
compilers. Finally Section~\ref{sec:Conclusion} summarizes our conclusions and
gives an outlook on further research topics.

@@

\section{Guards in the Frontend} %{Resume Data}
\label{sec:Resume Data}

@@

\section{Related Work}
\label{sec:Related Work}

@@

% subsection Deoptimization in Method-Based JITs (end)
% section Related Work (end)

\section{Conclusion}
\label{sec:Conclusion}
